### 2017-10-12 writing/paper/sections/coincidences.tex
Near final fixes

\begin{proposition}[Third Dominating Pattern Rule]
    \label{prop:dom3}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\), then
    \(m_1 \coinc{\pi} m_2\) if
    \begin{enumerate}
        \item The mesh set \(R_2 = R_1 \cup \{\boks{a,b}\}\)
        \item\label{prop:dom3:condocc} One of the patterns in \(m_1^\boksall{a,b}\)
            is coincident with a mesh pattern containing an occurrence of
            \(m_2\) as a subpattern.
    \end{enumerate}
\end{proposition}
\begin{proof}
    \NoteRef{not:downcmesh} implies \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\) as before.
    Now consider a permutation \(\omega\) in \(\av{\pi}\) that contains an
    occurrence of \(m_1\). If the region corresponding to the box \(\boks{a,b}\)
    is empty then we have an occurrence of \(m_2\).

    By condition~\ref{prop:dom3:condocc} of the proposition there exists
    a pattern \(r\) in \(m_1^\boksall{a,b}\) such that \(m_2\) is a subpattern in some
    pattern \(r'\) coincident with \(r\). If the region is non-empty then there exists
    an occurrence of \(r\) in \(\omega\). Therefore there is an occurrence of \(r'\)
    in \(\omega\), which implies that there is an occurrence of \(m_2\) in \(\omega\).
    Thus \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\) and the
    two patterns are coincident.
\end{proof}

This rule reduces the number of classes in
\(\av{\{\perm{2,3,1},(\perm{1,2}, R)\}}\) to \(56\), which coincides with the
56 classes observed through experimentation.

\subsection{Coincidence classes of Av(\{321, (12, \textit{R})\}).}
When considering coincidences of mesh patterns with underlying classical
pattern \(\perm{1,2}\) in \(\av{\perm{3,2,1}}\) application of the previously
established rules give no coincidences. Through experimentation we discover
that there are \(7\) non-trivial coincidence classes (all others are singletons)
which can be explained through the
use of two different lines of reasoning. Since the number of coincidences is
so small we will reason for these coincidences without attempting to generalise
into concrete rules.

Intuitively it is easy to see why our previous rules have no
power here. It is impossible to add a single point to a mesh pattern \((12, R)\) and
create an occurrence of \(\pi = \perm{3,2,1}\). It is also impossible to have
a position where addition of an ascent, or descent, provides extra
shading power.

The patterns
\begin{equation*}
    m_1 = \pattern{}{1,2}{0/0, 0/2, 1/1, 1/2, 2/1} \text { and } m_2 \pattern{}{1,2}{0/0, 0/1, 0/2, 1/1, 1/2, 2/1}
\end{equation*}
are coincident in \(\av{\perm{3,2,1}}\). (There are 3 symmetries of these
patterns that are also coincident to each other by the same reasoning.)

Consider the region corresponding to the box \((0,1)\) in any occurrence of \(m_1\),
in a permutation. By \LemmaRef{lem:incdec} it must contain an increasing subsequence.
If the region is empty then we have an occurrence of \(m_2\).
If there is only one point in the region we can choose this to replace the \(1\) in
the mesh pattern to get the required shading. If there is more than one point then
choosing the two leftmost points gives us the following mesh pattern.
\begin{equation*}
    \pattern{3,4}{2,3,1,4}{0/0, 0/1, 0/2, 0/3, 0/4,
                            1/0, 1/1, 1/2, 1/3, 1/4,
                            2/0, 2/1,2/2, 2/4,
                            3/1, 3/2, 3/3, 3/4,
                            4/1, 4/2, 4/3}
\end{equation*}
Where the two highlighted points are the original two points.
The other two points are an occurrence of the pattern
we originally desired, and hence the two patterns are coincident.
It is also possible to calculate this coincidence by an extension of the Third
Dominating rule, where we allow a sequence of point addition operations, this
is discussed further in the future work section.

The other reasoning applies to the patterns
\begin{equation*}
    m_1 = \pattern{}{1,2}{0/1,0/2,2/0,2/1} \text{ and } m_2 = \pattern{}{1,2}{0/1,0/2,1/1,2/0,2/1}
\end{equation*}
which are coincident by experimentation.

In order to prove this coincidence we will proceed by mathematical
induction on the number of points in region corresponding to the middle box.
We call this number \(n\).
\begin{description}
    \item [Base Case \((n=0)\)] The base case holds since we can freely shade the
        box if it contains no points.
    \item [Inductive Hypothesis \((n=k)\)] Suppose that we can find an occurrence
        of the second pattern if we have an occurrence of the first with \(k\) points
        in the middle box.
    \item [Inductive Step \((n=k+1)\)] Suppose that we have \((k+1)\) points in
        the middle box. Choose the bottom most point in the middle box, giving the
        mesh pattern
     \begin{equation*}
        \begin{tikzpicture}[scale=0.5]
            \modpattern[5]{}{1,2,3}{0/1,0/2,0/3,1/1,2/1,3/0,3/1,3/2}
            \draw (2.5,0.5) node {\(X\)};
        \end{tikzpicture}
    \end{equation*}
        Now we need to consider the box labelled \(X\). If this box is empty then
        we have an occurrence of \(m_2\) and are done. If this box contains
        any points then we gain some extra shading on the mesh pattern
        as any points in these boxes would create an occurrence of the dominating pattern \(\perm{3,2,1}\)
     \begin{equation*}
        \begin{tikzpicture}[scale=0.5]
            \modpattern[5]{2,3}{1,2,3}{0/1,0/2,0/3,1/1,1/2,1/3,2/1,3/0,3/1,3/2}
            \filldraw (2.5,0.5) circle (4 pt);
        \end{tikzpicture}
    \end{equation*}
        The two highlighted points form an occurrence of \(m_1\) with \(k\)
        points in the middle box, and thus by the Inductive Hypothesis we
        are done.
\end{description}
By induction we have that every occurrence of \(m_1\) leads to an occurrence of \(m_2\)
and by \NoteRef{not:downcmesh} every occurrence of \(m_2\) is an occurrence of
\(m_1\) so the two patterns are coincident.
This argument applies to another two pairs of classes. Therefore in
total in \(\av{\{\perm{3,2,1}, \mperm{1,2}{R}\}}\) there are \(213\) coincidence
classes.

>>>>

\begin{proposition}[Third Dominating Pattern Rule]
    \label{prop:dom3}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\), then
    \(m_1 \coinc{\pi} m_2\) if
    \begin{enumerate}
        \item The mesh set \(R_2 = R_1 \cup \{\boks{a,b}\}\)
        \item\label{prop:dom3:condocc} One of the patterns in \(m_1^\boksall{a,b}\)
            is coincident with a mesh pattern containing an occurrence of
            \(m_2\) as a subpattern.
    \end{enumerate}
\end{proposition}
\begin{proof}
    \NoteRef{not:downcmesh} implies \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\) as before.
    Now consider a permutation \(\omega \in \av{\pi}\) that contains an
    occurrence of \(m_1\). If the region corresponding to the box \(\boks{a,b}\)
    is empty then we have an occurrence of \(m_2\).
    By condition~\ref{prop:dom3:condocc} of the proposition there exists
    a pattern \(r\) in \(m_1^\boksall{a,b}\) such that \(m_2\) is a subpattern in some
    pattern \(r'\) coincident with \(r\). If the region is non-empty then there exists
    an occurrence of \(r\) in \(\omega\). Therefore there is an occurrence of \(r'\)
    in \(\omega\), which implies that there is an occurrence of \(m_2\) in \(\omega\).
    Thus \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\) and the
    two patterns are coincident.
\end{proof}

After implementing the Third Dominating Pattern Rule we find that there are
\(56\) \(\coinc[(3)]{231}\)-equivalence classes, of mesh patterns where the underlying
classical pattern is \(12\). By \NoteRef{note:main} there are
therefore exactly \(56\) \(\coinc{231}\)-equivalence classes, since we have
observed \(56\) \(\coinc[comp]{231}\)-equivalence classes through experimentation.

\subsection{Coincidence classes of Av(\{321, (12, \textit{R})\}).}
When considering coincidences of mesh patterns with underlying classical
pattern \(\perm{1,2}\) in \(\av{\perm{3,2,1}}\) application of the previously
established rules gives no coincidences. Through experimentation we discover
that there are \(7\) \(\coinc[comp]{321}\)-equivalence classes are unexplained.
Since the number of coincidences is
so small we will reason for these coincidences without attempting to generalise
into concrete rules.

Intuitively, it is easy to see why our previous rules have no
power here. It is impossible to add a single point to a mesh pattern \((12, R)\) and
create an occurrence of \(\pi = \perm{3,2,1}\). It is also impossible to have
a position where addition of an ascent, or descent, provides extra
shading power.

The patterns
\begin{equation*}
    m_1 = \pattern{}{1,2}{0/0, 0/2, 1/1, 1/2, 2/1} \text { and } m_2 \pattern{}{1,2}{0/0, 0/1, 0/2, 1/1, 1/2, 2/1}
\end{equation*}
are \(\coinc[comp]{321}\)-equivalent. (There are 3 symmetries of these
patterns that are also equivalent to each other by the same reasoning.)

Consider the region corresponding to the box \(\boks{0,1}\) in any occurrence of \(m_1\)
in a permutation. By \LemmaRef{lem:incdec} it must contain an increasing subsequence.
If the region is empty then we have an occurrence of \(m_2\).
If there is only one point in the region we can choose this point to replace the \(1\) in
the mesh pattern to get the required shading. If there is more than one point then
choosing the two leftmost points gives us the following mesh pattern.
\begin{equation*}
    \pattern{3,4}{2,3,1,4}{0/0, 0/1, 0/2, 0/3, 0/4,
                            1/0, 1/1, 1/2, 1/3, 1/4,
                            2/0, 2/1,2/2, 2/4,
                            3/1, 3/2, 3/3, 3/4,
                            4/1, 4/2, 4/3}
\end{equation*}
Here the two highlighted points are the original two points.
The other two points are an occurrence of the pattern
we \(m_2\), and hence the two patterns are coincident.
It is also possible to calculate this coincidence by an extension of the Third
Dominating rule, where we allow a sequence of point addition operations, this
is discussed further in the future work section.
Because of symmetries we have dealt with \(4\) out of the \(7\) \(\coinc[comp]{321}\)-equivalence classes.

Now consider the patterns
\begin{equation*}
    m_1 = \pattern{}{1,2}{0/1,0/2,2/0,2/1} \text{ and } m_2 = \pattern{}{1,2}{0/1,0/2,1/1,2/0,2/1}
\end{equation*}
which are \(\coinc[comp]{321}\)-equivalent.

In order to prove this coincidence we will proceed by mathematical
induction on the number of points in the region corresponding to the middle box.
We call this number \(n\).
\begin{description}
    \item [Base Case \((n=0)\)] The base case holds since we can freely shade the
        box if it contains no points.
    \item [Inductive Hypothesis \((n=k)\)] Suppose that we can find an occurrence
        of the second pattern if we have an occurrence of the first with \(k\) points
        in the middle box.
    \item [Inductive Step \((n=k+1)\)] Suppose that we have \((k+1)\) points in
        the middle box. Choose the bottom most point in the middle box, giving the
        mesh pattern
     \begin{equation*}
        \begin{tikzpicture}[scale=0.5]
            \modpattern[5]{}{1,2,3}{0/1,0/2,0/3,1/1,2/1,3/0,3/1,3/2}
            \draw (2.5,0.5) node {\(X\)};
        \end{tikzpicture}
    \end{equation*}
        Now we need to consider the box labelled \(X\). If this box is empty then
        we have an occurrence of \(m_2\) and are done. If this box contains
        any points then we gain some extra shading on the mesh pattern
        as any points in the boxes \(\boks{1,2}\) and \(\boks{1,3}\) would create an occurrence of the dominating pattern \(\perm{3,2,1}\)
     \begin{equation*}
        \begin{tikzpicture}[scale=0.5]
            \modpattern[5]{2,3}{1,2,3}{0/1,0/2,0/3,1/1,1/2,1/3,2/1,3/0,3/1,3/2}
            \filldraw (2.5,0.5) circle (4 pt);
        \end{tikzpicture}
    \end{equation*}
        The two highlighted points form an occurrence of \(m_1\) with \(k\)
        points in the middle box, and thus by the Inductive Hypothesis we
        are done.
\end{description}
By induction we have that every occurrence of \(m_1\) leads to an occurrence of \(m_2\)
and by \NoteRef{not:downcmesh} every occurrence of \(m_2\) is an occurrence of
\(m_1\) so the two patterns are coincident.
This argument applies to another two pairs of classes. Therefore in
total in \(\av{\{\perm{3,2,1}, \mperm{1,2}{R}\}}\) there are \(\coinc{321}\)-equivalence
classes.

### 2017-10-12 writing/paper/sections/wilf.tex
Near final fixes

Since we are always considering Wilf-equivalences in a set \(\av{S}\) we
must only use symmetries that preserve the dominating pattern(s), if we
were to allow other symmetries, then the equivalences calculated in the
previous section do not necessarily hold.

@@

The elements to the left of the maximum, \(A\), have the structure of a \(231\)
avoiding permutation, and the elements to the right of the maximum, \(B\), have
the structure of a \(\perm{2,3,1}\) avoiding permutation. Furthermore, all the
elements in \(A\) lie below all of the elements in \(B\). We call \(A\) the
\emph{lower-left section} and \(B\) the \emph{upper-right section}.

@@

The pattern \(\textpattern{}{1}{0/0,1/0,0/1,1/1}\) is in its own
Wilf-class since the only permutation containing this pattern is
the permutation \(\perm{1}\). The avoiders of this pattern therefore
have generating function \(\var(x) = C(x)-x\).

@@

\nextvar[\varfine]
It can be shown by use of \PropositionRef{prop:dom2} that the patterns
\textpattern{}{1}{0/0,1/1} and \(q_1 = \textpattern{}{1}{0/0,0/1,1/1}\) are
coincident inside \(\av{\perm{2,3,1}}\). Consider the decomposition of a non-empty avoider of \(q_1\) in
\(\av{\perm{2,3,1}}\) around the maximum:
\begin{equation*}
    \scriptvar_1 =\varepsilon \sqcup
\decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
\end{equation*}
This can be explained succinctly by the fact that a permutation containing
\(q_1\) starts with its maximum, and by not allowing the lower-left section of the
\(\perm{2,3,1}\) avoider to be empty we prevent an occurrence from ever
happening.

@@

\section{Wilf-classes with patterns of length 2}
By the use of coincidence classes established in
\SectionRef{sec:coincs} we know that there are at most 95 Wilf-equivalence classes.

@@

Computing avoiders up to length 10 suggests that there are at least 23 Wilf-classes,
of which 13 are non-trivial. Therefore there are Wilf-equivalences that are not
explained by coincidences or symmetry.

We will use two main methods of establishing Wilf-equivalence between mesh patterns of length \(2\)
in \(\av{\perm{2,3,1}}\): the structural decomposition of avoiders, via generating functions; or the
structure of a general permutation containing the pattern, looking at a particular occurrence of the
pattern in a permutation avoiding \(\perm{2,3,1}\). Sometimes it will be necessary to use both of
these methods to consolidate a single Wilf-class.

@@

If there is an occurrence of \(m_1\) there must be a lowest occurrence of \(m_1\), \ie the occurrence has the lowest possible values for any occurrence. Now consider the top
regions, labelled \(A \et B\), the subpermutation contained in these regions must avoid the permutation \(231\). Also, the subpermutation in the region \(A\) must be a decreasing subsequence, otherwise an occurrence of \(231\) will be created with either of the points in the occurrence. Now, consider the region labelled \(C\), since we specified that we were focused on the lowest possible occurrence of \(m_1\) this region cannot contain an occurrence of either \(m_1\) or \(231\). We now have a structural decomposition of a container of \(m_1\) inside \(\av{231}\).

>>>>

Since we are always considering Wilf-equivalences in a set \(\av{S}\) we
must only use symmetries that preserve the dominating pattern(s) in \(S\), if we
were to allow other symmetries, then the equivalences calculated in the
previous section do not necessarily hold.

@@

The elements to the left of the maximum, \(A\), have the structure of a \(231\)
avoiding permutation, and the elements to the right of the maximum, \(B\), also have
the structure of a \(\perm{2,3,1}\) avoiding permutation. Furthermore, all the
elements in \(A\) lie below all of the elements in \(B\). We call \(A\) the
\emph{lower-left section} and \(B\) the \emph{upper-right section}.

@@

The pattern \(\textpattern{}{1}{0/0,1/0,0/1,1/1}\) is in its own
Wilf-class since the only permutation containing it is
the permutation \(\perm{1}\). The avoiders of this pattern therefore
have generating function \(\var(x) = C(x)-x\).

@@

\nextvar[\varfine]
It can be shown by use of the Second Dominating Pattern Rule that the patterns
\textpattern{}{1}{0/0,1/1} and \(q_1 = \textpattern{}{1}{0/0,0/1,1/1}\) are
coincident under \(\perm{2,3,1}\). Consider the decomposition of a non-empty avoider of \(q_1\) in
\(\av{\perm{2,3,1}}\) around the maximum:
\begin{equation*}
    \scriptvar_1 =\varepsilon \sqcup
\decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
\end{equation*}
This can be explained succinctly by the fact that a permutation containing
\(q_1\) starts with its maximum, and by not allowing the lower-left section of the
\(\perm{2,3,1}\) avoider to be empty we prevent an occurrence from ever
happening.

@@

\section{Wilf-classes with patterns of length 2}
From \SectionRef{sec:coincs} we know there are \(95\) coincidence classes and
therefore at most that many Wilf-equivalence classes.

@@

Computing avoiders up to length 10 suggests that there are at least 23 Wilf-classes,
of which 13 are not equal to coincidence classes (after taking symmetries). Therefore there seem to be Wilf-equivalences that are not
explained by coincidences or symmetry.

We will use two main methods of establishing these Wilf-equivalences: the structural decomposition of avoiders, via generating functions; or the
structure of a general permutation containing the pattern, looking at a particular occurrence of the
pattern in a permutation avoiding \(\perm{2,3,1}\). Sometimes it will be necessary to use both of
these methods to consolidate a single Wilf-class.

@@

If there is an occurrence of \(m_1\) there must be a lowest occurrence of \(m_1\), \ie, the occurrence has the lowest possible values for any occurrence. Now consider the top
regions, labelled \(A \et B\), the subpermutation contained in these regions must avoid the permutation \(231\). Also, the subpermutation in the region \(A\) must be a decreasing subsequence, otherwise an occurrence of \(231\) will be created with either of the points in the occurrence. Now, consider the region labelled \(C\), since we specified that we were focused on the lowest possible occurrence of \(m_1\) this region cannot contain an occurrence of either \(m_1\) or \(231\). We now have a structural decomposition of a container of \(m_1\) inside \(\av{231}\).
